# IST707 Assignment 1 - nrangnan #
Shiny app link: https://nrangnan.shinyapps.io/Nisha/
``` {r}
attritionDF <- read.csv("employee_attrition.csv")
library(ggplot2)
library(dplyr)
library(arules)
library(arulesViz)
```
Structure of the data
```{r}
str(attritionDF)
```

We have total 9 NA's in the overall dataset
```{r}
sum(is.na(attritionDF))
```



Use Summary() to view details about each attribute in the dataset, such as Min, Max, Median, etc. for quantitative variables;
and frequency of each term for categorical variables.
It also tells you the number of NA's in each column
```{r}
summary(attritionDF) 
```


# Data Cleaning & preprocessing #

1. We can see from summary() output that DistanceFromHome attribute has 2 NA's.
To deal with these, we can safely replace these NA's by median of distance from home of all employees.
```{r}
attritionDF <- attritionDF %>%
  mutate(DistanceFromHome = replace(DistanceFromHome, is.na(DistanceFromHome), median(DistanceFromHome, na.rm=TRUE)))
```

2.JobLevel - We have 1 NA in the JobLevel column.
we can set job level as -1 so that it doesn't impact the support % when we use apriori.
```{r}
attritionDF <- attritionDF %>%
  mutate(JobLevel = replace(JobLevel, is.na(JobLevel), -1))
```

3. PercentSalaryHike - we have 1 NA.
Replace it with median of the attribute.
```{r}
attritionDF <- attritionDF %>%
  mutate(PercentSalaryHike = replace(PercentSalaryHike, is.na(PercentSalaryHike), median(PercentSalaryHike, na.rm=TRUE)))
```

4.PerformanceRating - We have 1 NA in the PerformanceRating column
we can set rating as -1 so that it doesn't impact the support % when we use apriori
```{r}
attritionDF <- attritionDF %>%
  mutate(PerformanceRating = replace(PerformanceRating, is.na(PerformanceRating), -1))
```

4.RelationshipSatisfaction - We have 1 NA in the RelationshipSatisfaction column
we can set satisfaction rating as -1 so that it doesn't impact the support % when we use apriori
```{r}
attritionDF <- attritionDF %>%
  mutate(RelationshipSatisfaction = replace(RelationshipSatisfaction, is.na(RelationshipSatisfaction), -1))
```

5. TotalWorkingYears - 2 NAs
You can replace it with their YearsInCurrentRole (assuming that they have been in this role always).
```{r}
attritionDF <- attritionDF %>%
  mutate(TotalWorkingYears = replace(TotalWorkingYears, is.na(TotalWorkingYears), YearsInCurrentRole[which(is.na(TotalWorkingYears))]))
```

6. Years Since Last Promotion -
we can set it as -1 so that it doesn't impact the support % when we use apriori.
```{r}
attritionDF <- attritionDF %>%
  mutate(YearsSinceLastPromotion = replace(YearsSinceLastPromotion, is.na(YearsSinceLastPromotion), -1))
```

We dont have any more NA's
```{r}
sum(is.na(attritionDF))
```

Lets remove duplicates, if any
```{r}
attritionDF <- attritionDF %>%
  distinct()
dim(attritionDF) 
```
There were no duplicate rows (as dimensions of the data remain the same).


# Exploratory Data Analysis #

Explore Attrition rate
```{r}
g <- ggplot(attritionDF, aes(x=attritionDF$Attrition))
g <- g + geom_bar()
g
```

1. Age
```{r}
gg1 <- ggplot(attritionDF, aes(x=attritionDF$Attrition, y=attritionDF$Age)) + geom_boxplot()
gg1
```

2. BusinessTravel
```{r}
g2 <- ggplot(attritionDF, aes(x=attritionDF$BusinessTravel, fill = attritionDF$Attrition))
g2 <- g2 + geom_bar()
g2
```

3. Department wise attrition
```{r}
g3 <- ggplot(attritionDF, aes(x=attritionDF$Department, fill= attritionDF$Attrition))
g3 <- g3 + geom_bar(position = "dodge")
g3
```

4. Distance from home
```{r}
gg4 <- ggplot(attritionDF, aes(x=attritionDF$Attrition, y=attritionDF$DistanceFromHome)) + geom_boxplot()
gg4
```

5. EducationLevel
```{r}
g5 <- ggplot(attritionDF, aes(x=attritionDF$Education, fill= attritionDF$Attrition))
g5 <- g5 + geom_bar(position = "dodge")
g5
```

6. EducationField
```{r}
g6 <- ggplot(attritionDF, aes(x=attritionDF$EducationField, fill= attritionDF$Attrition))
g6 <- g6 + geom_bar(position = "dodge")
g6
```

7. Environment Satisfaction
```{r}
g7 <- ggplot(attritionDF, aes(x=attritionDF$EnvironmentSatisfaction, fill= attritionDF$Attrition))
g7 <- g7 + geom_bar(position = "dodge")
g7
```

8. Gender
```{r}
g8 <- ggplot(attritionDF, aes(x=attritionDF$Gender, fill= attritionDF$Attrition))
g8 <- g8 + geom_bar(position = "dodge")
g8
```

9. JobInvolvement
```{r}
g9 <- ggplot(attritionDF, aes(x=attritionDF$JobInvolvement, fill= attritionDF$Attrition))
g9 <- g9 + geom_bar(position = "dodge")
g9
```
The percentage for job attrition is the least amongst employees that are highly involved in their job

10. JobLevel
```
g10 <- ggplot(attritionDF, aes(x=attritionDF$JobLevel, fill= attritionDF$Attrition))
g10 <- g10 + geom_bar(position = "dodge")
g10
```

11. JobRole
```{r}
g11 <- ggplot(attritionDF, aes(x=attritionDF$JobRole, fill= attritionDF$Attrition))
g11 <- g11 + geom_bar(position = "dodge")
g11 <- g11 + theme(axis.text.x = element_text(angle = 60))
g11
```

12. JobSatisfaction
```{r}
g12 <- ggplot(attritionDF, aes(x=attritionDF$JobSatisfaction, fill= attritionDF$Attrition))
g12 <- g12 + geom_bar(position = "dodge")
g12
```
The more satisfied you are with your job, the less likely you are to leave it

13. Marital Status
```{r}
g13 <- ggplot(attritionDF, aes(x=attritionDF$MaritalStatus, fill= attritionDF$Attrition))
g13 <- g13 + geom_bar(position = "dodge")
g13
```

14. MonthlyIncome
```{r}
gg14 <- ggplot(attritionDF, aes(x=attritionDF$Attrition, y=attritionDF$MonthlyIncome)) + geom_boxplot()
gg14
```

15. NumCompaniesWorked
```{r}
g15 <- ggplot(attritionDF, aes(x=attritionDF$NumCompaniesWorked, fill= attritionDF$Attrition))
g15 <- g15 + geom_bar(position = "dodge")
g15
```

16. oVERTIME
```{r}
g16 <- ggplot(attritionDF, aes(x=attritionDF$OverTime, fill= attritionDF$Attrition))
g16 <- g16 + geom_bar(position = "dodge")
g16
```
People who work overtime are more likely to leave

17. PerformanceRating
```{r}
g17 <- ggplot(attritionDF, aes(x=attritionDF$PerformanceRating, fill= attritionDF$Attrition))
g17 <- g17 + geom_bar(position = "dodge")
g17
```

18. Relationship satisfaction
```{r}
g18 <- ggplot(attritionDF, aes(x=attritionDF$RelationshipSatisfaction, fill= attritionDF$Attrition))
g18 <- g18 + geom_bar(position = "dodge")
g18
```

19. Total working years
```{r}
gg19 <- ggplot(attritionDF, aes(x=attritionDF$Attrition, y=attritionDF$TotalWorkingYears)) + geom_boxplot()
gg19
```

20. WorkLifeBalance
```{r}
g20 <- ggplot(attritionDF, aes(x=attritionDF$WorkLifeBalance, fill= attritionDF$Attrition))
g20 <- g20 + geom_bar(position = "dodge")
g20
```
Percentage of attrition goes down as the level of work-life balance goes up

21. YearsAtCompany
```{r}
gg21 <- ggplot(attritionDF, aes(x=attritionDF$Attrition, y=attritionDF$YearsAtCompany)) + geom_boxplot()
gg21
```

22. Years since last promotion
```{r}
gg22 <- ggplot(attritionDF, aes(x=attritionDF$Attrition, y=attritionDF$YearsSinceLastPromotion)) + geom_boxplot()
gg22
```

23Years with current manager
```{r}
gg23 <- ggplot(attritionDF, aes(x=attritionDF$Attrition, y=attritionDF$YearsWithCurrManager)) + geom_boxplot()
gg23
```


# Data transformation - Deleting columns with redundant data, conversion of certain column data to factors #
```{r}
attritionDF$Education <- as.factor(attritionDF$Education)
attDF1 <- attritionDF[,-9]

attDF1$JobInvolvement <- as.factor(attDF1$JobInvolvement)
attDF1$PerformanceRating <- as.factor(attDF1$PerformanceRating)

attDF1 <- attDF1[,-26]

attDF1$StockOptionLevel <- as.factor(attDF1$StockOptionLevel) 
attDF1$WorkLifeBalance <- as.factor(attDF1$WorkLifeBalance)

attDF1 <- attDF1[,-21]
```

# Association rules mining & tuning of hyperparameters  #
## Rules for Attrition = No ##
```{r}
rules <- apriori(attDF1, parameter = list(supp = 0.5, conf = 0.8, target = "rules", minlen=2), appearance = list(rhs="Attrition=No"))
rules
rules_lift <- sort(rules, by = "lift", decreasing = TRUE)
inspect(head(rules_lift))
a <- plot(rules)
print(a)
```

Decrease number of rules by increasing support value
```{r}
rules <- apriori(attDF1, parameter = list(supp = 0.56, conf = 0.8, target = "rules", minlen=2), appearance = list(rhs="Attrition=No"))
rules
rules_lift <- sort(rules, by = "lift", decreasing = TRUE)
inspect(head(rules_lift))
a <- plot(rules)
print(a)
``` 

## Rules for Attrition = Yes ##
```{r}
rules1 <- apriori(attDF1, parameter = list(supp = 0.2, conf = 0.5, target = "rules", minlen=2), appearance = list(rhs="Attrition=Yes"))
rules1
```
received 0 rules, decrease support value

```{r}
rules1 <- apriori(attDF1, parameter = list(supp = 0.043, conf = 0.5, target = "rules", minlen=2), appearance = list(rhs="Attrition=Yes"))
rules1
rules_lift1 <- sort(rules1, by = "lift", decreasing = TRUE)
inspect(head(rules_lift1))
a <- plot(rules1)
print(a)
```


# Interpretation of attrition=Yes #
 Employees are likely to leave their job if they're not at a high job position, paid low monthly income and they work overtime.
Employees are more likely to leave if they're a young working crowd, at a job level 1 or 2, with not too high work experience (0 to 6 years), having low number of years ar the company; 
i.e. if they have not been at the company for very long, and they have been promoted recently, they are likely to switch the job to get a better position and salary.
Employees that have low montly income, who work overtime are more likely to leave the job.

# Interpretation of attrition=No #
Senior position employees are less likely to leave the company.
Employees who do not work overtime are less likely to leave the company.
Research & Development employees are less likely to leave the company.
Employees that are not required to travel often (travel_rarely) are less likely to leave the company.
